---
title: "ESM 244 Lab 1"
author: "Grace Brofman"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#installing package (do it in the console)
#install.packages("packagename") e.g. "sf" and "tmap"
library(tidyverse)
library(here)
library(sf)
library(tmap)

```

### Read in SF trees data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))

```

### Basic wrangling reminders

Refresh some skills for data wrangling and summary statistics using functions in the `dplyr` package.

Find the top 5 highest observations of trees by legal_status, do some wrangling, make a graph.
count function
1. recognizes groups (replaces group_by)
2. finds counts (replaces n function)
3. puts in table (replaces summarize)

drop_na removes any row that contains missing or na value for the variable that you specify
rename renames a column (new_name = old_name)
relocate: allows you to move things around by class, location, or single column
lice_max allows you to identify rows with highest values that youve identified and keep those rows
```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n = 5)
  

```

Make a graph of those top 5 observations by legal status.
```{r}
ggplot(data = top_5_status,
       aes(x = fct_reorder(legal_status, tree_count),
           y = tree_count)) +
  geom_col() +
  labs(x = "Legal Status",
       y = "Tree Count") +
  coord_flip() +
  theme_minimal()

# default is to plot in alphabetical order. we'd prefer them from high to low. change legal status to a factor that is associated with tree count

```

### A few more data wrangling refresher examples

Only want to keep observations (rows) for Blackwood Acacia trees.

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude) 

#making a map using lat/long as spatial coordinates
ggplot(data= blackwood_acacia,
       aes(x = longitude,
           y = latitude)) +
  geom_point()


```

### tidyr::separate() and unit() functions

Useful for combining or separating columns

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")

```

Example: tidyr::unite()

combining tree and legal status columns into single new column with separator that you identify
```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_")
```

### Make some actual maps of blackwood acacia trees in SF

`st_as_sf()` to convert latitude and longitute to spatial coordinates.

```{r}
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(longitude, latitude) %>% 
  st_as_sf(coords = c("longitude", "latitude")) #list whatever variable names store long/lat

#set CRS (coordinate reference system) if not assigned (4326 is 4 digit code that is common)
st_crs(blackwood_acacia_sp) = 4326

#geom_sf() is for plotting spacial data in ggplot (as opposed to geom_point())
ggplot(data = blackwood_acacia_sp) +
  geom_sf(color = "darkgreen")

```

REad in SF roads shapefile:
```{r}
sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

#need to make sure that these are in the same coord reference system (CRS). in this case there's an exitins crs so we need to use st_transform()

st_transform(sf_map, 4326)

ggplot(data = sf_map) +
  geom_sf()

```

Combine blackwood acacia tree observations and SF roads map:

```{r}
ggplot() +
  geom_sf(data = sf_map, 
          size = 0.1,
          color = "darkgray") +
  geom_sf(data = blackwood_acacia_sp, 
          color = "red",
          size = 0.5) +
  theme_void()
```

Now an interactive map:
```{r}
tmap_mode("view")

tm_shape(blackwood_acacia_sp) +
  tm_dots()

```

