---
title: "ESM 244 Lab 1"
author: "Grace Brofman"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#installing package (do it in the console)
#install.packages("packagename") e.g. "sf" and "tmap"
library(tidyverse)
library(here)
library(sf)
library(tmap)

```

### Read in SF trees data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))

```

### Basic wrangling reminders

Refresh some skills for data wrangling and summary statistics using functions in the `dplyr` package.

Find the top 5 highest observations of trees by legal_status, do some wrangling, make a graph.
count function
1. recognizes groups (replaces group_by)
2. finds counts (replaces n function)
3. puts in table (replaces summarize)

drop_na removes any row that contains missing or na value for the variable that you specify
rename renames a column (new_name = old_name)
relocate: allows you to move things around by class, location, or single column
lice_max allows you to identify rows with highest values that youve identified and keep those rows
```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n = 5)
  

```








